---
title: "Registered students for HBO 2014 - 2018"
author: "Marc A.T. Teunis"
date: "`r Sys.Date()`"
params:
  croho: "techniek"
  institute: "Hogeschool Utrecht"
  education_track: "B Biologie en Medisch Laboratoriumonderzoek" 
  education_type: "bachelor"
output:
  BiocStyle::html_document:
    toc: true
    toc_depth: 2
---

# Packages
```{r}
library(tidyverse)
library(here)
library(naniar)
library(ggridges)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.width = 10, fig.height = 7)
theme(legend.position = "none")
```

# Data

## Data - GET
```{r}
path_to_datafile <- here("data-raw", "D010", "01-ingeschrevenen-hbo-2018.csv")
data <- read_csv2(path_to_datafile)
```

## Data - INSPECT
```{r}
head(data)
head(data[, c("2014 Man", "2018 Vrouw")])
```

## Data - TIDY
```{r}
data_tidy <- data %>%
  gather(`2014 Man`:`2018 Vrouw`,
         key = "year_gender", value = "counts") %>%
  separate(year_gender, into = c("year", "gender"),
           remove = FALSE)
data_tidy[, c("year", "gender", "counts")]
```

## Data - FIX NAMES
```{r}
names(data_tidy) <- names(data_tidy) %>%
  str_replace_all(pattern = " ", replacement = "_") %>%
  tolower()
names(data_tidy)
```

## Data - MISSING VALUES
For this analysis we remove all NAs
```{r}
sum(is.na(data_tidy))
vis_miss(data_tidy)
naniar::gg_miss_var(data_tidy)
naniar::gg_miss_upset(data_tidy)
data_tidy <- na.omit(data_tidy)

```

# Exploratory Data Analysis

## Distributions

- Many zero or low counts
- The median is `r median(data_tidy$counts, na.rm = TRUE)`
```{r}
median <- median(data_tidy$counts, na.rm = TRUE) %>% log10() 
hist(data_tidy$counts)
data_tidy %>%
  mutate(counts = counts + 0.001) %>% ## add a pseudocount of 0.001 for log10 transform
  ggplot(aes(x = log10(counts))) +
  geom_freqpoly() +
  geom_vline(xintercept = median, colour = "darkred", linetype = "dashed", size = 1)
```

## Ditrubutions per chroho
```{r}
data_tidy %>%
   mutate(counts = counts + 0.001) %>% ## add a pseudocount of 0.001 for log10 transform
  na.omit() %>%
  ggplot(
    aes(x = log10(counts),
    y = croho_onderdeel,
    fill = croho_onderdeel)) +
      geom_density_ridges() +
      theme(legend.position = "none")+
      labs(title = "Distributions per CROHO") +
  facet_wrap(~ year)
```

## All crohos over time
```{r}
data_tidy %>%
#  dplyr::filter(counts < 10) %>%
  group_by(year, croho_onderdeel) %>%
  summarise(sum_counts = sum(counts)) %>%
  ggplot(aes(x = year, y = sum_counts)) +
  geom_point(aes(colour = croho_onderdeel,
                 shape = croho_onderdeel), size = 2) +
  geom_line(aes(colour = croho_onderdeel, 
                group = croho_onderdeel), size = 1) +
  scale_shape_manual(values = c(1:8)) +
  theme_bw()
```

## General pattern over time for all institutes, cummulative
Here we examine the registered students on the global institutional level. All enrollments are summed per institute and year to get a general overview.
```{r}
library(toolboxr)
data_tidy %>%
  group_by(instellingsnaam_actueel) %>%
  summarise(sum_counts = sum(counts)) %>%
  ggplot(aes(x = sum_counts, 
             y = reorder(
               as_factor(instellingsnaam_actueel), sum_counts
               ))) +
  geom_point() +
  theme_bw() +
#  facet_wrap(~year, nrow = 1) +
  ylab("Institutes") +
  xlab("Sum counts") +
  rotate_axis_labels(axis = "x", angle = 90) +
  scale_x_continuous(labels = scales::scientific)
```

## General pattern over time for all institutes, over the years
Here we examine the registered students on the global institutional level. All enrollments are summed per institute and year to get a general overview.
```{r}
library(toolboxr)
data_tidy %>%
  group_by(year, instellingsnaam_actueel) %>%
  summarise(sum_counts = sum(counts)) %>%
  ggplot(aes(x = sum_counts, 
             y = reorder(
               as_factor(instellingsnaam_actueel), sum_counts
               ))) +
  geom_point() +
  theme_bw() +
  facet_wrap(~year, nrow = 1) +
  ylab("Institutes") +
  xlab("Sum counts") +
  rotate_axis_labels(axis = "x", angle = 90) +
  scale_x_continuous(labels = scales::scientific)
```

# Parameterized report
To zoom in on one institute, the remainder of this report is paramterized for one institute, croho or education track

## Generic pattern for education tracks within one croho, only the larger educational tracks (> 3000 total enrollments, in each year)
```{r}
library(directlabels)
library(ggrepel)

data_tidy %>%
  dplyr::filter(croho_onderdeel == params$croho) %>%
  group_by(opleidingsnaam_actueel, year) %>%
 
  summarise(sum_counts = sum(counts)) %>%
  filter(sum_counts > 3000) %>%
  mutate(label = if_else(year == max(year), 
                         as.character(
                           opleidingsnaam_actueel), 
                         NA_character_)) %>%
  
  
    ggplot(aes(x = year, y = sum_counts)) +
  geom_point(aes(colour = opleidingsnaam_actueel), size = 1.5) +
  geom_line(aes(colour = opleidingsnaam_actueel, 
                group = opleidingsnaam_actueel), size = 1) +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle(paste("All enrollments for the croho \n",
                params$croho, ", for all education tracks within this croho", sep = "")) +
  
  scale_x_discrete(expand=c(0, 2)) +
  geom_label_repel(aes(label = label),
                  nudge_x = 3,
                  nudge_y = 2,
                  na.rm = TRUE)



data_tidy %>%
  dplyr::filter(croho_onderdeel == params$croho) %>%
  group_by(opleidingsnaam_actueel, year) %>%
 
  summarise(sum_counts = sum(counts)) -> highest

high_track <- highest %>%
  filter(sum_counts == max(highest$sum_counts))

```

## Educational with maximum enrollment in croho `r params$croho` track `r high_track$opleidingsnaam_actueel` at a closer look
```{r}
data_tidy %>%
  dplyr::filter(croho_onderdeel == params$croho) %>%
  group_by(opleidingsnaam_actueel, year) %>%
  summarise(sum_counts = sum(counts)) %>%
  
  mutate(label = if_else(year == max(year), 
                         as.character(
                           opleidingsnaam_actueel), 
                         NA_character_)) %>%
  
  dplyr::filter(opleidingsnaam_actueel %in% high_track$opleidingsnaam_actueel) %>%
  ggplot(aes(x = year, y = sum_counts)) +
  geom_line(aes(colour = opleidingsnaam_actueel, 
                group = opleidingsnaam_actueel), size = 3) +
  
  geom_point(size = 3, colour = "blue") + 
  theme_bw() +
  theme(legend.position = "none") +
  
  scale_x_discrete(expand=c(0, 1.3)) +
  ylim(c(10000, 20000)) +
  geom_label_repel(aes(label = label),
                  nudge_x = 3,
                  nudge_y = 2,
                  na.rm = TRUE) 



```


## Generic pattern for `r params$education_track`
```{r}
data_tidy %>%
  dplyr::filter(opleidingsnaam_actueel == params$education_track) %>%
  group_by(opleidingsnaam_actueel, year) %>%
  summarise(sum_counts = sum(counts)) %>%
  ggplot(aes(x = year, y = sum_counts)) +
  geom_point(aes(colour = opleidingsnaam_actueel), size = 2, colour = "blue") +
  geom_line(aes(colour = opleidingsnaam_actueel, 
                group = opleidingsnaam_actueel), size = 1,
            colour = "blue") +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle(paste("Number of enrollments for \n", 
                params$education_track)) +
  ylim(4000, 6000)
```

### Generic pattern for `r params$education_track` over institutes providing this track.
```{r}
data_tidy %>%
  dplyr::filter(opleidingsnaam_actueel == params$education_track) %>%
  group_by(opleidingsnaam_actueel, year, instellingsnaam_actueel) %>%
  summarise(sum_counts = sum(counts)) %>%
  ggplot(aes(x = year, y = sum_counts)) +
  geom_point(aes(colour = opleidingsnaam_actueel), 
             colour = "darkblue") +
  geom_line(aes(colour = opleidingsnaam_actueel, 
                group = opleidingsnaam_actueel), 
            colour = "darkblue") +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle(paste("Number of total enrollments for \n", 
                params$education_track, "per institute")) +
  facet_wrap(~ instellingsnaam_actueel)
```

# Gender differences

## For croho `r params$croho`
```{r}
data_tidy %>%
  dplyr::filter(croho_onderdeel == params$croho) %>%
#  mutate(counts_log2 = log2(counts + 0.0001)) %>%
#  dplyr::filter(counts_log2 > 2.5) %>%
  ggplot(aes(x = year,
             y = log10(counts))) +
  geom_point(aes(colour = gender),
             shape = 1, position = "jitter") +
  geom_smooth(method = "lm") +
  
  ggtitle(paste("Man vrouw verdeling bij croho \n",
                params$croho)) +
  theme_bw()
```

## For educational track `r params$education_track`
Male female distributions
```{r}
data_tidy %>%
  dplyr::filter(
    opleidingsnaam_actueel == params$education_track) %>%
  mutate(counts_log2 = log2(counts + 0.0001)) %>%
  dplyr::filter(counts_log2 > 2.5) %>%
  ggplot(aes(x = year,
             y = counts_log2)) +
  geom_point(aes(colour = gender),
             shape = 1, position = "jitter") +
  geom_smooth(aes(group = gender, colour = gender), method = "loess", se = FALSE) +
  ggtitle(paste("Man vrouw verdeling bij \n",
                params$education_track)) +
  theme_bw()
```

## Overall differences in registered females vs males for `r params$educational_track`
```{r}
data_tidy %>%
  dplyr::filter(
    opleidingsnaam_actueel == params$education_track) %>%
#  mutate(counts_log2 = log2(counts + 0.0001)) %>%
  group_by(gender, year) %>%
  summarise(sum_counts = sum(counts)) %>%
#  dplyr::filter(counts_log2 > 2.5) %>%
  ggplot(aes(x = year,
             y = sum_counts)) +
  geom_point(aes(colour = gender)) +
  geom_smooth(aes(group = gender, colour = gender)) +
#  geom_smooth(aes(group = gender, colour = gender), method = "loess", se = FALSE) +
  scale_color_manual(values = c("blue", "red")) +
  theme_bw()
```

## Education types, comparisons
Here we compare Bachelor vs Deeltijd and Master educations, we look at differences between males and females per croho, for the institute: `r params$institute



